---
title: "STATS 326/786"
author: "Week 8 Tutorial"
date: \today
fontsize: 11pt
output:
  bookdown::html_document2:
    fig_height: 5
    toc: yes
    toc_depth: 1
    toc_float:
      collapsed: false
    number_sections: false
    code_folding: show
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
library(tidyverse)
library(fpp3)
library(grid)
library(gridExtra)
```


# Problem 1

Consider the number of pigs slaughtered in Victoria, available in the `aus_livestock` data set.  This is plotted below.

```{r, echo = FALSE}
pigs <- aus_livestock %>%
  filter(State == "Victoria") %>%
  filter(Animal == "Pigs")

pigs %>% 
  autoplot(Count) + 
  theme_minimal()
```


> Fit three different simple exponential smoothing (`ETS(A,N,N)`) models.  One with $\alpha = 0.001$, one with $\alpha = 0.999$ and one with $\alpha$ that is automatically optimised.  You can specify the value of $\alpha$ in the `trend` function.  

```{r}

```

> For the model where $\alpha$ was automatically optimised, extract and interpret the estimates of $l_0$ and $\alpha$.  

```{r}

```

> On three seperate plots overlay the fitted values on the original time series.  Comment on what you observe.  

```{r}

```


> Using the `weight.plot` function provided below, plot the weights that the past 20 observations contribute to the exponenentially smoothed value at the current time point.  Comment on what you observe. 

```{r}

# Function to use
weight.plot = function(alpha, N = 20) {
  
  # alpha is the smoothing parameter for the level
  # N is the number of time points in the past to plot
  
  Weight = alpha    # First weight
  for (i in 2:N) {  # Other weights
    Weight[i] = Weight[i - 1] * (1 - alpha)  
  }
  
  # Create tsibble
  data <- as_tsibble(data.frame(Time = (-(1:N)),
                      Weight = Weight), 
           index = Time)

  # Plot
  ggplot(data = data, mapping = aes(x = Time, y = Weight)) +
    geom_segment(aes(x = Time, xend = Time,
                     y = 0, yend = Weight)) +
    theme_bw()

}

```

> For all three models, forecast 4 years into the future ($h = 48$) and plot these forecasts with their default prediction intervals. Comment on what you observe.

```{r}

```


# Extra: Problem 2

Consider the Algerian exports time series in the `global_economy` data set.  This is plotted below.  An ETS(A,N,N) model is fitted to it and the optimal parameters are reported.  

```{r}
algeria_economy <- global_economy %>%
  filter(Country == "Algeria") 

algeria_economy %>%
  autoplot(Exports) +
  theme_bw()

fit <- algeria_economy %>%
  model(ETS(Exports ~ error("A") + trend("N") + season("N")))
report(fit)  # Extract parameters
```


> Write your own function to implement simple exponential smoothing. The function should take arguments `y` (the time series), `alpha` (the smoothing parameter, $\alpha$) and `level` (the initial level, $l_0$). It should return the fitted values of the exponentially smoothed time series. Check to see that this function produces the same fitted values as `ETS()`.

> Modify your function from the previous exercise to return the sum of squared errors rather than the fitted values. Then use the `optim` function to find the optimal values of $\alpha$ and $l_0$. Do you get the same values as the `ETS()` function?

```{r}

```
