---
title: "STATS 326/786"
author: "Week 1 Tutorial"
date: \today
fontsize: 11pt
output:
  bookdown::html_document2:
    fig_height: 5
    toc: yes
    toc_depth: 1
    toc_float:
      collapsed: false
    number_sections: false
    code_folding: show
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
library(tidyverse)
```



> The National Basketball Association (NBA) is North America's premier men's professional basketball league.  Data has been collected on 54 random NBA players.  The following variables have been measured: height (in feet), weight (in pounds), fgp (field goal percentage), ftp (free-throw percentage), and ppg (average points-per-game).

> 1. Read in `basketball.csv` using `read_csv` and convert the data frame to a `tibble` object.

```{r, message = FALSE}

```

> 2. Use the `mutate` function to create the following new variables in your `tibble` object: height (in cm), weight (in kg), body-mass-index (BMI), and a binary indicator that determines if an NBA player is tall or short (where a player is classified as tall if they are taller than 6.5 ft).

The following conversions will help:

\begin{eqnarray*}
\text{height (m)} &=& \text{height (ft)} \times 0.3048 \\
\text{weight (kg)} &=& \text{weight (lb)} \times 0.45359237 \\
\text{bmi} &=& \frac{\text{weight (kg)}}{\text{height (m)}^2}
\end{eqnarray*}

Hint: The `ifelse` function may be useful for creating your binary indicator.

```{r}

```

> 3. Use the `group_by` and `summarise` functions to calculate the mean field goal percentage, free throw percentage, and points-per-game for short versus tall NBA players.  

```{r}

```

> 4. Find the largest and smallest average points-per-game for tall versus short NBA players.  

```{r}

```

> 5. Use `ggplot` and `geom_point` to make a scatter plot of weight (in kg) by height (in m), and size by BMI. Use the `alpha` command add transparency.  Comment on the plot.  

```{r}

```

> 6. Use `ggplot` and `geom_point` to make a scatter plot of field goal percentage by average points-per-game, with short and tall NBA players on different panels.  Comment on the plot.  

```{r}

```

> 7. Use `ggplot` and `geom_histogram` to make a histogram of average points-per-game, colouring by your binary height indicator.  Use the `alpha` command to add transparency.  Comment on the plot.

```{r}

```


> 8. If you finish early, play around with other `geom` functions, `aes` mappings, as well as other `scale` and `theme` options.

```{r}

```