---
title: "STATS 326/786"
author: "Week 4 Tutorial"
date: \today
fontsize: 11pt
output:
  bookdown::html_document2:
    fig_height: 5
    toc: yes
    toc_depth: 1
    toc_float:
      collapsed: false
    number_sections: false
    code_folding: show
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
library(fpp3)
```


# Problem 1

> Below is some fake seasonal data (with a perfectly periodic pattern) with no trend and no remainder components.  In class, we learned that moving averages are useful for smoothing time series. Compute a $2 \times 4~\text{MA}$ and a $5~\text{MA}$ and overlay these on the original time series.  What do you notice, and why is this happening?  

```{r, warning = FALSE}

data <- tsibble(Quarter = yearquarter(1:40),
                y = rep(c(4, 9, 3, 7), 10),
                index = Quarter)

data %>% 
  autoplot(y, colour = "grey") + 
  theme_minimal()

```


> Repeat the above exercise, but for the fake data below.  This data is seasonal (again, with a perfectly periodic pattern) with a linear trend and no remainder component.  

```{r, warning=FALSE}

data <- tsibble(Quarter = yearquarter(1:40),
                y = rep(c(4, 9, 3, 7), 10) + 0.5 * 1:40,
                index = Quarter)

data %>% 
  autoplot(y, colour = "grey") + 
  theme_minimal()

```


> If you were to apply a $3 \times 5~\text{MA}$ to a time series, analytically derive the weights that each observation contributes to the moving-averaged time series at time $t$.  



# Problem 2

> Below is an example of a time series with multiplicative seasonality.  In this problem, you learn how to perform your own (multiplicative) classical decomposition from first principles.  

```{r}
a10 <- PBS %>%
  filter(ATC2 == "A10") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(Total_Cost = sum(Cost)) %>%
  mutate(Total_Cost = Total_Cost / 1e6,
         Month.Name = month(Month, label = TRUE)) 

a10 %>%
  autoplot(Total_Cost) + 
  theme_minimal()
```

> Calculate the trend-cycle by performing a $2\times 12~\text{MA}$ on your time series.

```{r}

```
         
> Compute the detrended time series (original time series divided by trend).    
         
```{r}         

```

> Calculate the seasonal factor for each month (noting that these must sum to 12 for a monthly time series).  

```{r}

```

> Merge your data sets so the seasonal factor is now included in your original `tsibble`.

```{r}

```

> Compute the remainder component (as detrended divided by seasonal) and seasonally-adjusted component (as original time series divided by seasonal).

```{r}

```

> Plot the original time series and overlay it with the trend-cycle and seasonally-adjusted series.  

```{r, warning = FALSE}

```

> Extra: See if you can re-create the classical decomposition plot.

```{r, warning = FALSE}

```
