---
title: "STATS 326/786"
author: "Week 2 Tutorial"
date: \today
fontsize: 11pt
output:
  bookdown::html_document2:
    fig_height: 5
    toc: yes
    toc_depth: 1
    toc_float:
      collapsed: false
    number_sections: false
    code_folding: show
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
library(lubridate)
library(tidyverse)
```

# Problem 1

> Use `lubridate` functions to calculate how many days old you are as of today. Then calculate how long you've been alive in years.

```{r}

```


# Problem 2

> Matt went to a conference in Toulouse in 2022.  He departed Auckland at 1515 on 18 November 2022, and landed in Toulouse, France at 1220 on 19 November.  Use `lubridate` functions to compute the number of hours he was in transit for.  Be careful with time-zones.

```{r}

```


# Problem 3

> Thanksgiving Day in the United States is the fourth Thursday of November.  Find the date for Thanksgiving Day in 2025.

```{r}

```


# Problem 4

> Write an `R` function that outputs yesterday's date, and write an `R` function that outputs tomorrow's date.

```{r}

```


# Problem 5

> The `lubridate` package has functions periods, such as `seconds`, `weeks`, and `years`.  There is no function in `lubridate` for fortnights.  Write your own `fortnights` `R` function using pre-existing `lubridate` functions. Run your function and output the periods for 1, 2, and 3 fortnights.  

```{r}

```

# Problem 6

> The `flights` data set in the `nycflights13` package contains information on all flights from New York in 2013.  Consider all flights from JFK (US/Eastern time-zone) to LAX (US/Pacific time-zone).  Install and call this `R` package and type `?flights` to look at the help file to see variable details.

```{r, echo = FALSE}
library(nycflights13)

# ?flights  # Look at help file to see variable details

data = flights %>%
  filter(origin == "JFK",
         dest == "LAX",
         !is.na(arr_time)) 
```

## (a) 

> Create a departure date-time variable by considering the actual departure time and departure delay.  Note that `time_hour` is the date-time variable for the scheduled departure time.  

```{r}

```


## (b)

> Create an arrival date-time variable by considering the actual arrival time, the date of departure, and whether a flight arrives on the same day or the following day (overnight flights). 

```{r}

```


## (c)

> We are told that the average transit time is 6.5 hours.  Create a variable to calculate the transit time, i.e., the difference between the arrival date-time and departure date-time.  Plot the distribution and comment. Then plot the transit time against the air time and comment.  

```{r, message = FALSE, warning = FALSE}

```
